---
title: "Peernorms"
author: "Emmy Green"
date: "11/29/2018"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(ggplot2)
library(dplyr)
library(forcats)
library(lme4)
library(devtools)
library(MASS)
library(nnet)
library(citr)
library(RefManageR)
library(survival)
library(survminer)
library(car)
library(brms)
library(plyr)
library(effects)
library(broom)
library(purrr)
library(car)
library(gee)
library(geepack)
library(lavaan)
library(psych)

```

#Calling the Data
```{r,include=FALSE}

#Call from choose
peernorms<-read.csv(file.choose())

```

```{r, include=FALSE}

#Data frames

#Drop kids
peernorms<-peernorms%>% 
  filter(drop==0)

#Define age_exact as numeric variable
peernorms$age_exact<-as.numeric(peernorms$age_exact)

#Define gender as categorical variable
peernorms$gender <- as.factor(peernorms$gender)

#Define race as categorical variable
peernorms$race<-factor(peernorms$race, levels = c(0, 1))

#Relabel race cateogry labels
peernorms$race<-revalue(peernorms$race, c("0" = "White", "1" = "Minority"))

#Define condition as categorical
peernorms$condition <-factor(peernorms$condition, levels = c("0", "1"))

#Relabel condition labels
peernorms$condition <-revalue(peernorms$condition, c("0" = "control", "1" = "friendship"))

#Defining neighborhood as numeric 
peernorms$neighborhoodbg<-as.numeric(peernorms$neighborhoodbg)
peernorms$neighborhoodbb<-as.numeric(peernorms$neighborhoodbb)
peernorms$neighborhoodwg<-as.numeric(peernorms$neighborhoodwg)
peernorms$neighborhoodwb<-as.numeric(peernorms$neighborhoodwb)

### Creating composite variables

#Attitudes towards black target (rescaled to be 1-6; higher means positive about black)
peernorms<-peernorms%>%
  mutate(attitude_black = (likebg + likebb + nicebg + nicebb)/4)

#Attitudes towards White target (rescaled to be 1-6; higher means positive about white)
peernorms<-peernorms%>%
  mutate(attitude_white = (likewg + likewb + nicewg + nicewb)/4) 

#ProWhite bias (scale: -5 to 5; positive number means bias towards whites)
peernorms<-peernorms%>%
  mutate(pro_white_bias = (attitude_white - attitude_black))

#ProWhite bias - GENDER INGROUP (scale: -10 to 10; positive number means bias towards whites)
peernorms<-peernorms%>%
  mutate(prowhitebias_gingroup = ((likeingroupw + niceingroupw) - (likeingroupb + niceingroupb)))

#ProWhite bias - GENDER OUTGROUP (scale: -10 to 10; positive number means bias towards whites)
peernorms<-peernorms%>%
  mutate(prowhitebias_goutgroup = ((likeoutgroupw + niceoutgroupw) - (likeoutgroupb + niceoutgroupb)))

#Creating Neighborhood Diversity (scale: higher means more likely to see black than white)
peernorms<-peernorms%>%
  mutate(neighborhood = (neighborhoodbg + neighborhoodbb) - (neighborhoodwg + neighborhoodwb))

#Creating similar variable
peernorms<- peernorms%>%
  mutate(similar= (similarities1 + similarities2 + similarities3 + similarities4)/4)

#Creating the expectations for helpful behavior variable:
peernorms<- peernorms%>%
  mutate(expectations_helpful= (expectations_white + expectations_black)/2)

#Creating the peer norms variable:
peernorms<- peernorms%>%
  mutate(norms= (norms_white +norms_black)/2)

#Set Neighborhood Diveresity as numeric variable
peernorms$neighborhood<-as.numeric(peernorms$neighborhood)

#Centering age_exact
peernorms<-peernorms%>%
  mutate(age_exactc = (age_exact - (mean(age_exact))))

#Centering Neighborhood Diversity
peernorms$neighborhoodc<-scale(peernorms$neighborhoodc,center=TRUE,scale=FALSE)

```

#Analyses

### Attitudes Towards Blacks Measure
```{r, include=FALSE}

##Descriptives
#Comparing means by condition

aggregate(peernorms$similar~peernorms$condition,FUN=mean)
aggregate(peernorms$attitude_black~peernorms$condition,FUN=mean)
aggregate(peernorms$expectations_friend~peernorms$condition,FUN=mean)
aggregate(peernorms$expectations_black~peernorms$condition,FUN=mean)
aggregate(peernorms$expectations_white~peernorms$condition,FUN=mean)
aggregate(peernorms$norms~peernorms$condition,FUN=mean)

aggregate(cbind(peernorms$similar,peernorms$attitude_black,peernorms$expectations_friend) ~ peernorms$condition, FUN=mean) 

#LM ATB
attitudemod<-lm(attitude_black ~ age_exactc*race, data = peernorms)
attitudemod_summary<-as.data.frame(coef(summary(attitudemod)))

summary (attitude_black)

#ANOVA
Anova(attitude_black)

#ATB and condition (friendship or control)
attitudemod2<-lm(attitude_black~age_exactc*race*condition, data= peernorms)
attitudemod2_summary <-as.data.frame(coef(summary(attitudemod2)))

#ANOVA
Anova(attitudemod2)

##Figure attitudes
attitudesfigure1<-ggplot(peernorms, aes(x=condition, y=attitude_black, color = race)) +
  geom_jitter(shape = 1) +
  geom_smooth (method = "glm", alpha = .1) +
  ylim(1, 6) +
  scale_color_manual(values = c("red", "blue", "purple"))+
  theme_bw() +
  labs(x="Condition", y="Children's Attitudes Towards the Black Targets")

attitudesfigure1

```

### Pro-White Bias
```{r, include=FALSE}

#LM Pro White Bias
pro_white_mod<-lm(pro_white_bias ~ age_exactc*race, data = peernorms)
pro_white_mod_summary<-as.data.frame(coef(summary(pro_white_mod)))

summary (pro_white_mod)

#ANOVA
Anova(pro_white_mod)

#Pro white and condition
pro_white_mod2<-lm(pro_white_bias ~ age_exactc*race*condition, data = peernorms)
pro_white_mod2_summary<-as.data.frame(coef(summary(pro_white_mod2)))

##Figure pro white bias
prowhitefigure<-ggplot(peernorms, aes(x=condition, y=attitude_white, color = race)) +
  geom_jitter(shape = 1) +
  geom_smooth (method = "glm", alpha = .1) +
  ylim(1, 6) +
  scale_color_manual(values = c("red", "blue", "purple"))+
  theme_bw() +
  labs(x="Condition", y="Children's Pro-White Bias")

prowhitefigure

#expectations
expectations_friend_mod<- lm(expectations_friend ~ age_exactc*race*condition, data=peernorms)
expectations_friend_mod_summary<-as.data.frame(coef(summary(expectations_mod1)))

expectations_white_mod<- lm(expectations_white ~ age_exactc*race*condition, data=peernorms)
expectations_white_mod_summary<-as.data.frame(coef(summary(expectations_white_mod)))

expectations_black_mod<- lm(expectations_black ~ age_exactc*race*condition, data=peernorms)
expectations_black_mod_summary<-as.data.frame(coef(summary(expectations_black_mod)))

expectations_helpful_mod<- lm(expectations_helpful~ age_exactc*race*condition, data= peernorms)
expectations_helpful_summary<-as.data.frame(coef(summary(expectations_helpful_mod)))

#similarity

similar_mod <- lm(similar ~ age_exactc*race*condition, data=peernorms)
similar_mod_summary<-as.data.frame(coef(summary(similar_mod)))

#norms
norms_white_mod <- lm(norms_white ~ age_exactc*race*condition, data=peernorms)
norms_white_mod_summary<-as.data.frame(coef(summary(norms_white_mod)))

norms_black_mod <- lm(norms_black ~ age_exactc*race*condition, data=peernorms)
norms_black_mod_summary<-as.data.frame(coef(summary(norms_black_mod)))

norms_mod <- lm(norms ~ age_exactc*race*condition, data=peernorms)
norms_mod_summary<-as.data.frame(coef(summary(norms_mod)))

#Mediation analyses 

#Expectations of friendship as a mediator 

peernorms<-peernorms
na.omit(peernorms,cols=c(“condition”,“expectations_friend”,“attitude_black”))

sem1<- “# a path
expectations_friend ~ a*condition  #mediator as a function of IV

 #b path
attitude_black ~ b*expectations_friend  #DV as a function of mediator

 #c’path
attitude_black ~ cp*condition  #c-prime path aka DV as function of IV

 ab := a*b
 total := cp + ab”

set.seed(1234)

sem1<-sem(sem1,data=peernorms,se=“bootstrap”,bootstrap = 1000)
summary(sem1,standardized=TRUE)

#Expectations of helpful behavior as a mediator 

peernorms<-peernorms
na.omit(peernorms,cols=c(“condition”,“expectations_helpful”,“attitude_black”))

sem2<- “# a path
expectations_helpful ~ a*condition  #mediator as a function of IV

 #b path
attitude_black ~ b*expectations_helpful  #DV as a function of mediator

 #c’path
attitude_black ~ cp*condition  #c-prime path aka DV as function of IV

 ab := a*b
 total := cp + ab”

set.seed(1234)

sem2<-sem(sem1,data=peernorms,se=“bootstrap”,bootstrap = 1000)
summary(sem1,standardized=TRUE)

#Expectations of perceived similarity as a mediator 

peernorms<-peernorms
na.omit(peernorms,cols=c(“condition”,“similar”,“attitude_black”))

sem3<- “# a path
similar ~ a*condition  #mediator as a function of IV

 #b path
attitude_black ~ b*similar  #DV as a function of mediator

 #c’path
attitude_black ~ cp*condition  #c-prime path aka DV as function of IV

 ab := a*b
 total := cp + ab”

set.seed(1234)

sem3<-sem(sem1,data=peernorms,se=“bootstrap”,bootstrap = 1000)
summary(sem1,standardized=TRUE)

#Expectations of peer norms as a mediator 

peernorms<-peernorms
na.omit(peernorms,cols=c(“condition”,“norms”,“attitude_black”))

sem4<- “# a path
norms~ a*condition  #mediator as a function of IV

 #b path
attitude_black ~ b*norms  #DV as a function of mediator

 #c’path
attitude_black ~ cp*condition  #c-prime path aka DV as function of IV

 ab := a*b
 total := cp + ab”

set.seed(1234)

sem4<-sem(sem1,data=peernorms,se=“bootstrap”,bootstrap = 1000)
summary(sem1,standardized=TRUE)

