---
title: "Children remain interested and feeling efficacious about their capacity to 'do science,' but lose confidence in their potential to 'be scientists'"
author: "Ryan F. Lei, Emmy Green, & Marjorie Rhodes"
date: "6/15/2018"
output:
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load libraries needed for these analyses (any uninstalled packages will need to be installed).

library(lmerTest)
library(lme4)
library(tidyr)
library(effects)
library(ggplot2)
library(psych)
library(dplyr)
library(tidyverse)
library(knitr)
library(mediation)
library(RColorBrewer)

```
*Longitudinal Results*
##contrast coding the data##
```{r, include = FALSE}
#Read in the data file
#This data file contains data from all three waves

setwd("/Volumes/Server/Data/Studies/Sci S (Longitudinal Science)/Clean data for analysis")
science<-read.csv("Longitudinal_long_6.15.18.csv")
science2 <- science%>% 
  filter(drop==0)
sci_graph<-science2
## Frequencies 
N=xtabs(~ time + gender, science2)
N
N2=xtabs(~time, science2)
N2

##contrast coding variables

science2$gender[science2$gender == 1] <- -.5
science2$gender[science2$gender == 2] <- .5

science2$time[science2$time == 1] <- -1
science2$time[science2$time == 2] <- 0
science2$time[science2$time == 3] <- 1

science2<-science2%>%
  filter(condition == "Be" | condition == "Do")
science2$condition<-as.character(science2$condition)
science2$condition[science2$condition == "Be"] <- -.5
science2$condition[science2$condition == "Do"] <- .5
science2$condition<-as.numeric(science2$condition)

##Interest in Science##

#Recoding 2 to 0 in the binominal question (do you want to be a scientist/do science? Yes/No)
science2$yousci[science2$yousci==2] <- 0

sciprev<-science2%>%
  filter(time!=-1)

##re-coding for graph

sci_graph$time[sci_graph$time == 1] <- -1
sci_graph$time[sci_graph$time == 2] <- 0
sci_graph$time[sci_graph$time == 3] <- 1

```

## HELPER FUNCTIONS FOR WITHIN SUBJECTS ERROR BARS##
```{r,include=FALSE}

## HELPER FUNCTIONS FOR WITHIN SUBJECTS ERROR BARS##

normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
                           na.rm=FALSE, .drop=TRUE) {
  library(plyr)
  
  # Measure var on left, idvar + between vars on right of formula.
  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
                         .fun = function(xx, col, na.rm) {
                           c(subjMean = mean(xx[,col], na.rm=na.rm))
                         },
                         measurevar,
                         na.rm
  )
  
  # Put the subject means with original data
  data <- merge(data, data.subjMean)
  
  # Get the normalized data in a new column
  measureNormedVar <- paste(measurevar, "_norm", sep="")
  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
    mean(data[,measurevar], na.rm=na.rm)
  
  # Remove this subject mean column
  data$subjMean <- NULL
  
  return(data)
}

summarySEwithin <- function(data=NULL, measurevar, betweenvars=NULL, withinvars=NULL,
                            idvar=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
  
  # Ensure that the betweenvars and withinvars are factors
  factorvars <- vapply(data[, c(betweenvars, withinvars), drop=FALSE],
                       FUN=is.factor, FUN.VALUE=logical(1))
  
  if (!all(factorvars)) {
    nonfactorvars <- names(factorvars)[!factorvars]
    message("Automatically converting the following non-factors to factors: ",
            paste(nonfactorvars, collapse = ", "))
    data[nonfactorvars] <- lapply(data[nonfactorvars], factor)
  }
  
  # Get the means from the un-normed data
  datac <- summarySE(data, measurevar, groupvars=c(betweenvars, withinvars),
                     na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
  
  # Drop all the unused columns (these will be calculated with normed data)
  datac$sd <- NULL
  datac$se <- NULL
  datac$ci <- NULL
  
  # Norm each subject's data
  ndata <- normDataWithin(data, idvar, measurevar, betweenvars, na.rm, .drop=.drop)
  
  # This is the name of the new column
  measurevar_n <- paste(measurevar, "_norm", sep="")
  
  # Collapse the normed data - now we can treat between and within vars the same
  ndatac <- summarySE(ndata, measurevar_n, groupvars=c(betweenvars, withinvars),
                      na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
  
  # Apply correction from Morey (2008) to the standard error and confidence interval
  #  Get the product of the number of conditions of within-S variables
  nWithinGroups    <- prod(vapply(ndatac[,withinvars, drop=FALSE], FUN=nlevels,
                                  FUN.VALUE=numeric(1)))
  correctionFactor <- sqrt( nWithinGroups / (nWithinGroups-1) )
  
  # Apply the correction factor
  ndatac$sd <- ndatac$sd * correctionFactor
  ndatac$se <- ndatac$se * correctionFactor
  ndatac$ci <- ndatac$ci * correctionFactor
  
  # Combine the un-normed means with the normed results
  merge(datac, ndatac)
}

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

##Main analyses##
```{r, include = FALSE}
#Analyses testing for effects of condition and time on whether children want to be a scientist/do science. This is a Generalized Linear Mixed Model (GLMM), using "glmer" to specify a binomial distribution. Random intercepts are included for participant and classroom.

choicemod<-glmer(cbind(yousci, 1-yousci) ~ condition*time 
                 + (1|id) + (1|classroom), 
                 data = science2, 
                 family = binomial,
                 glmerControl(optimizer = "bobyqa",
                optCtrl = list(maxfun=2e5)))
summary(choicemod)
tablechoicemod<-as.data.frame(coef(summary(choicemod)))
tablechoicemod

```
To investigate whether children were more likely to want to do science than be scientists, we analyzed their responses using a binomial multi-level model (see Figure 1A). Results revealed that overall, children more often wanted to do science (M = .85, SD = .14) than be a scientist (M= .65,*SD*=.13 ), $\beta$ = `r round((tablechoicemod[2,1]),2)`, *SE* = `r round((tablechoicemod[2,2]),2)`, *z* = `r round((tablechoicemod[2,3]),2)`, *p* = `r format.pval((tablechoicemod[2,4]), digits = 2, eps = .001)`.  Additionally, in general, children were less interested in science over time, $\beta$ = `r round((tablechoicemod[3,1]),2)`, *SE* = `r round((tablechoicemod[3,2]),2)`, *z* = `r round((tablechoicemod[3,3]),2)`, *p* = `r format.pval((tablechoicemod[3,4]), digits = 2, eps = .001)`. 
```{r, echo = FALSE, warning = FALSE}

## Plot effects of be/do condition and time on children's interest in science 

df_wc <- summarySEwithin(science2, measurevar="yousci", 
                         withinvars="time", 
                         betweenvars = "condition",
                        idvar="id", na.rm=TRUE, conf.interval=.95)
df_wc$time<-factor(df_wc$time,labels = c("Baseline", "4 Months", "8 Months"))
df_wc$condition<-factor(df_wc$condition,labels = c("Identity-Focused","Action-Focused"))
df_wc

attach(df_wc)
yousci_means<-aggregate(df_wc, by = list(condition), FUN = mean, na.rm=TRUE)
print(yousci_means)
detach(df_wc)

you_plot<-ggplot(df_wc, aes(x=time,y=yousci,group=1))+
  geom_point(data=df_wc, aes(x=time,y=yousci,
                             colour=condition,fill=condition),
             shape=21, size = 3)+
  geom_errorbar(width=.2, aes(ymin=yousci-ci,
                               ymax=yousci+ci))+
  ylim(0,1)+
  guides(color = guide_legend(reverse = TRUE)) +
                guides(fill = guide_legend(reverse = TRUE)) +
  labs(x="Time",y="Interest in Science")+
   #guides(color = guide_legend(reverse = TRUE))+
  theme_classic() + theme(text=element_text(size=14))
you_plot
```

```{r, include = FALSE}
#We used a Linear Mixed Model, using "lmer" to examine the effects of condition and time on how much children want to do science/be scientists. 

modelhowmuch<-lmer(howmuch~condition*time + (1|classroom) + (1|id), 
                   data = science2)
summary(modelhowmuch)
tablehowmuch<-as.data.frame(coef(summary(modelhowmuch)))
tablehowmuch


howmuch_graph<-lmer(howmuch~condition*time + (1|classroom) + (1|id), 
                   data = sci_graph)

meanhowmuch<-effect("condition*time", 
                    xlevels=list(condition=c(-.5,.5),
                                 time=c(-1,0,1)),
                    howmuch_graph)
meanhowmuch<-as.data.frame(meanhowmuch)
#meanhowmuch$time<-factor(meanhowmuch$time,labels = c("Wave 1", "Wave 2", "Wave 3"))
meanhowmuch$condition<-factor(meanhowmuch$condition,labels = c("Identity-Focused","Action-Focused"))
meanhowmuch

meanhowmuch2<-effect("condition",modelhowmuch)
meanhowmuch2<-as.data.frame(meanhowmuch2)
meanhowmuch2
```
We also examined whether children differed in how much they wanted to do science or be scientists (see Figure 1B). Overall, children wanted to do science more (M= 3.1, *SD*= .06) than they wanted to be scientists  (M= 2.8, *SD*= .05), $\beta$ = `r round((tablehowmuch[2,1]),2)`, *SE* = `r round((tablehowmuch[2,2]),2)`, *t* = `r round((tablehowmuch[2,4]),2)`, *p* = `r format.pval((tablehowmuch[2,5]), digits = 2, eps = .001)`, with no main or interactive effects of time. 
```{r, echo = FALSE, warning = FALSE}

## Plot effects of be/do condition and time on children's how much children wanted to do science/be scientists. 
View(sci_graph)

sci_graph$condition<-revalue(sci_graph$condition,c("Be"="Identity-focused",
                                             "Do"="Action-focused"))

plot_howmuch<-ggplot(data=sci_graph, aes(x=time, y=howmuch, colour=condition,linetype=condition)) + 
  geom_smooth(se=TRUE,method=lm,fullrange=TRUE)+
    geom_jitter(data=sci_graph,aes(x = time, y = howmuch, color= condition),
                 stat = "identity", size = 1.5, alpha = .3, 
                height = 0, width = 0.25, shape=1)+
  scale_x_discrete(limits = c(-1, 0, 1), labels = c("Baseline", "4 Months", "8 Months")) +
    labs(x= "Time", y="Extent of Interest")  + 
   scale_color_discrete(name= "Condition", labels = c("Identity-focused", "Action-focused")) +
  scale_linetype_discrete(name= "Condition", labels = c("Identity-focused", "Action-focused")) +
  guides(color = guide_legend(reverse = TRUE)) +
                guides(linetype = guide_legend(reverse = TRUE)) +
                theme_classic() + theme(text=element_text(size=14))
  
plot_howmuch
``` 

```{r, include=FALSE}

#We used a Linear Mixed Model, using "lmer" to examine the effects of condition and time on how good children thought they would be at doing science/being a scientist 

modelhowgood<-lmer(howgood~condition*time + (1|classroom) + (1|id), data=science2)
summary(modelhowgood)
tablehowgood<-as.data.frame(coef(summary(modelhowgood)))
tablehowgood

meanhowgood<-effect("condition*time", modelhowgood)
meanhowgood<-as.data.frame(meanhowgood)
meanhowgood

meanhowgood2<-effect("condition", modelhowgood)
meanhowgood2<-as.data.frame(meanhowgood2)
meanhowgood2
```
We next examined children’s self-efficacy in science (see Figure 2). Children thought they would be better at doing science (M=3.1, *SD* = .05)  than being scientists (M = 2.8, *SD* = .05), $\beta$ = `r round((tablehowgood[2,1]),2)`, *SE* = `r round((tablehowgood[2,2]),2)`, *t* = `r round((tablehowgood[2,4]),2)`, *p* = `r format.pval((tablehowgood[2,5]), digits = 2, eps = .001)`. Additionally, there was an effect of time such that children generally thought they would be worse at science over time, $\beta$ = `r round((tablehowgood[3,1]),2)`, *SE* = `r round((tablehowgood[3,2]),2)`, *t* = `r round((tablehowgood[3,4]),2)`, *p* = `r format.pval((tablehowgood[3,5]), digits = 2, eps = .001)`. Most importantly, these effects were subsumed by a significant interaction between condition and time, such that children in the identity-focused condition thought they would be worse at science over time, but children in the action-focused condition showed no significant change over time, $\beta$ = `r round((tablehowgood[4,1]),2)`, *SE* = `r round((tablehowgood[4,2]),2)`, *t* = `r round((tablehowgood[4,4]),2)`, *p* = `r format.pval((tablehowgood[4,5]), digits = 2, eps = .001)`.
```{r, echo = FALSE, warning = FALSE}

## Plot effects of be/do condition and time on children's self efficacy in science

plot_howgood <- ggplot(data=sci_graph, aes(x=time, y=howgood, colour=condition,linetype=condition)) + 
  geom_smooth(se=TRUE,method=lm,fullrange=TRUE)+
    geom_jitter(data=sci_graph,aes(x = time, y = howgood, color= condition),
                 stat = "identity", size = 1.5, alpha = .3, height = 0, width = 0.25, shape = 1)+
  scale_x_discrete(limits = c(-1, 0, 1), labels = c("Baseline", "4 Months", "8 Months")) +
    labs(x= "Time", y="Science Self-Efficacy")  + 
   scale_color_discrete(name= "Condition", labels = c("Identity-focused", "Action-focused")) +
  scale_linetype_discrete(name= "Condition", labels = c("Identity-focused", "Action-focused")) +
  guides(color = guide_legend(reverse = TRUE)) +
    guides(linetype = guide_legend(reverse = TRUE)) +
                theme_classic() + theme(text=element_text(size=14))
plot_howgood


```

```{r, include = FALSE}

## prevalence of scientists 

modelprev<-lmer(prevalence~condition*time + (1|classroom) + (1|id), data = sciprev)
summary(modelprev)
tableprev<-as.data.frame(coef(summary(modelprev)))
tableprev

meanprevalence<-effect("condition*time", modelprev)
meanprevalence<-as.data.frame(meanprevalence)
meanprevalence

meanprevalence2<-effect("condition", modelprev)
meanprevalence2<-as.data.frame(meanprevalence2)
meanprevalence2
```

##WITHIN-WAVE MEDIATION##
```{r, include= FALSE}

##MEDIATION ANALYSES##
##FYI -- lmerTest messes with the mediation package, so do not load it when conducting mediation analyses##

##within wave 2 mediation##

detach("package:lmerTest")

sciprev2<-sciprev %>%
  filter(time!=1)

sciprev2<-sciprev2 %>%
    filter(condition=="Be" | condition == "Do")
  
b<-lm(prevalence~condition,data=sciprev2)
c<-lm(howgood~condition+prevalence,data=sciprev2)
contcont<-mediate(b,c,sims=10000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont)
plot(contcont)


b1<-lm(prevalence~condition,data=sciprev2)
c1<-lm(howmuch~condition+prevalence,data=sciprev2)
contcont1<-mediate(b1,c1,sims=10000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont1)
plot(contcont1)


b2<-lm(prevalence~condition,data=sciprev2)
c2<-glm(yousci~condition+prevalence,data=sciprev2)
contcont2<-mediate(b2,c2,sims=10000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont2)

## mediation within wave 3 only##
sciprev3<-sciprev %>%
  filter(time==1)

View(sciprev3)
b_w3<-lm(prevalence~condition,data=sciprev3)
c_w3<-lm(howgood~condition+prevalence,data=sciprev3)
contcont_hg_w3<-mediate(b_w3,c_w3,sims=1000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont_hg_w3)
plot(contcont_hg_w3)

b2_w3<-lm(prevalence~condition,data=sciprev3)
c2_w3<-lm(howmuch~condition+prevalence,data=sciprev3)
contcont_hm_w3<-mediate(b2_w3,c2_w3,sims=1000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont_hm_w3)
plot(contcont_hm_w3)


b3_w3<-lm(prevalence~condition,data=sciprev3)
c3_w3<-glm(yousci~condition+prevalence,data=sciprev3)
contcont_yousci_w3<-mediate(b3_w3,c3_w3,sims=1000,
                  treat="condition",
                  mediator= "prevalence")
summary(contcont_yousci_w3)
plot(contcont_yousci_w3)

```

##longitudinal mediation##
```{r, include=FALSE}

##Longitudinal mediation##

##use sciprev as the df
sciprev_lm<-sciprev
sciprev_lm$wave<-sciprev_lm$time

sciprev_lm<-sciprev_lm[c(3,5,15,17,19,26:29,45)]

sciprev_lm$condition[sciprev_lm$condition == -.5] <- "Be"
sciprev_lm$condition[sciprev_lm$condition == .5] <- "Do"
sciprev_lm$condition<-as.factor(sciprev_lm$condition)
View(sciprev_lm)
#sciprev_lm<-sciprev_lm %>% 
#  filter(condition == "Be" | condition == "Do" )

med1<-lmer(prevalence~ condition*time + (1|id), data=sciprev_lm)
outcome.fit<-lmer(howgood~condition*prevalence*time + (1|id),data=sciprev_lm)

med.out_overall<-mediate(med1,outcome.fit,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 control.value = -.5,
                 treat.value = .5)
summary(med.out_overall)

med.out_t2<-mediate(med1,outcome.fit,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 0,
                 control.value = -.5,
                 treat.value = .5)
summary(med.out_t2)

med.out_t3<-mediate(med1,outcome.fit,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 1,
                 control.value = "Be",
                 treat.value = "Do")
summary(med.out_t3)
##mediation for the how much measure##

outcome.fit2<-lmer(howmuch~condition*prevalence*time + (1|id),data=sciprev_lm)

med.out2_overall<-mediate(med1,outcome.fit2,sims=1000,
                 treat="condition",
                 mediator="prevalence")
summary(med.out2_overall)

med.out2_t2<-mediate(med1,outcome.fit2,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 0)
summary(med.out2_t2)

med.out2_t3<-mediate(med1,outcome.fit2,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 1)
summary(med.out2_t3)

##mediation for the yousci measure##

outcome.fit3<-glmer(cbind(yousci,1-yousci)~condition*prevalence*time + (1|id),
                    family=binomial, data=sciprev_lm,
                     glmerControl(optimizer = "bobyqa",
                optCtrl = list(maxfun=2e5)))

med.out3_overall<-mediate(med1,outcome.fit3,sims=1000,
                 treat="condition",
                 mediator="prevalence")
summary(med.out3_overall)

med.out3_t2<-mediate(med1,outcome.fit3,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 0)
summary(med.out3_t2)

med.out3_t3<-mediate(med1,outcome.fit3,sims=1000,
                 treat="condition",
                 mediator="prevalence",
                 covariates = 1)
summary(med.out3_t3)

```

